Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-10-12 paper.tex
Add figures

\documentclass[draft,english]{jflart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lstcoq}
\usepackage[numbers]{natbib}
\usepackage{enumitem}

@@

  In defiance of Hinchliffe's rule, this article sets out to
  demonstrate that its title can be answered by the word ``yes''. We
  show that modern symmetric ciphers can be modeled through a small
  set of algebraic structures (Boolean algebra, Naperian and foldable
  functors). We reveal that these ciphers exhibit some interesting
  compositional structure at the type-level, which enables systematic
  code transformation, such as bitslicing and fixslicing. Our work
  rests on a Coq development providing the specification of two
  ciphers, Skinny and Gift, deriving a bitsliced and fixsliced
  implementation for each and proving their correctness with respect
  to their specifications.

@@

This article grew out of the Bachelor's research project of the first
author and the Master thesis of the second author. The former worked
on the specification, bitslicing and fixslicing of the Skinny
cipher~\citep{skinny-spec, adomin:skinny}. The latter worked on the
specification, bitslicing and fixslicing of the Gift
cipher~\citep{gift-spec, admon:gift}. Both ciphers are provided in the
aaccompanying source code~\footnote{\url{TODO}}. For clarity, we shall
focus exclusively on Skinny as our running example here.

@@

\begin{figure}
  %\includegraphics{skinny_state}
  \caption{Skinny state matrix}
  \label{fig:state}
\end{figure}

@@

\begin{figure}[htp]
  \caption{Shiftrows}
  \label{fig:shiftrows}

  \caption{Mixcolumns}
  \label{fig:mixcolumns}
\end{figure}

@@

\begin{figure}

  \caption{S8}
  \label{fig:s8}
\end{figure}

>>>>

\documentclass[draft,english]{jflart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lstcoq}
\usepackage[numbers]{natbib}
\usepackage{enumitem}
\usepackage{subcaption}

@@

  In defiance of Hinchliffe's rule, this article sets out to
  demonstrate that its title can be answered by the word ``yes''. We
  show that modern symmetric ciphers can be modeled through a small
  set of algebraic structures (Boolean algebra, Naperian and foldable
  functors). We reveal that these ciphers exhibit some interesting
  compositional structure at the type-level, which enables systematic
  code transformation, known in the cryptographical folklore as
  ``bitslicing'' and ``fixslicing''. Our work rests on a Coq
  development providing the specification of two ciphers, Skinny and
  Gift, deriving a bitsliced and fixsliced implementation for each and
  proving their correctness with respect to their specifications.

@@

This article grew out of the Bachelor's research project of the first
author and the Master thesis of the second author. The former worked
on the specification, bitslicing and fixslicing of the Skinny
cipher~\citep{skinny-spec, adomin:skinny}. The latter worked on the
specification, bitslicing and fixslicing of the Gift
cipher~\citep{gift-spec, admon:gift}. Both ciphers are provided in the
aaccompanying source code~\footnote{\url{https://github.com/pedagand/bitfix}}. For clarity, we shall
focus exclusively on Skinny as our running example here.

@@

\begin{figure}
  \begin{center}
    \includegraphics[draft=false,height=5cm]{skinny_state}
  \end{center}
  \caption{Skinny state matrix}
  \label{fig:state}
\end{figure}

@@

\begin{figure}[htp]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[draft=false,height=2.8cm]{shiftrows}
    \caption{Shiftrows (from~\citet{skinny-spec})}
    \label{fig:shiftrows}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[draft=false,height=2.8cm]{mixcolumns}
    \caption{Mixcolumns (from~\citet{skinny-spec})}
    \label{fig:mixcolumns}
  \end{subfigure}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[draft=false,height=5cm]{s8}
  \caption{Substitution box S8 (from~\citet{skinny-spec})}
  \label{fig:s8}
\end{figure}
